Portfolio conservation of metapopulations under climate change\\Climate change and portfolio conservation of metapopulations\\Climate change and portfolio conservation\\Portfolio prioritization of salmon conservation\\Portfolio prioritization of metapopulation conservation

### 2013-06-11 inst/ms/ms.tex
Develop the main text

\begin{itemize}
\item
  Portfolio optimization reveals\ldots{}
\item
  Portfolio theory reveals conservation strategies that\ldots{}
\item
  Intelligent tinkering with ecological portfolios: patterns of
  environmental response diversity drive salmon conservation priorities
\item
  Keeping every cog and wheel\ldots{}
\end{itemize}

@@

  \begin{itemize}
  \itemsep1pt\parskip0pt\parsep0pt
  \item
    risk is usually assessed on a population by population basis
  \item
    for metapopulations, we can consider risk for components or risk for
    the whole
  \end{itemize}
\item
  The management of financial portfolios provides another way of
  considering risk for metapopulations.

@@

Here, we ask how a portfolio approach to management can inform the
conservation of metapopulations in a changing world. We ask two primary
questions: (1) What does portfolio optimization tell us about different
spatial approaches to prioritizing metapopulation conservation assuming
that response diversity is spatially distributed? (2) In a scenario
where we don't know the population-level response diversity, what does
portfolio optimization tell us about the number of populations
conserved? To answer these questions, we develop a salmon metapopulation
simulation in which population-specific productivity is driven by
spatially-distributed environmental tolerance and patterns of short- and
long-term climatic change. We then implement different conservation
``rules of thumb'' that control the population-level carrying
capacities. We evaluate the salmon portfolios along risk and return
axes, similarly to how financial portfolios are assessed. Our findings
illustrate that: (1) conserving response diversity buffers
metapopulation risk given short-term climate forcing and rate of return
given long-term climate forcing, (2) conserving more subpopulations
buffers risk regardless of the climate trend, and (3) considering
metapopulations through the lens of portfolio theory provides a useful
additional dimension through which we can evaluate conservation
strategies.

@@

In our ecological portfolios, we defined assets as stream-level
populations and the portfolios as salmon metapopulations (Table
\ref{t:port}). We use the terms \textit{stream} and \textit{populations}
interchangeably to represent the portfolio assets. We defined the
portfolio investors as the stakeholders in the fishery and
metapopulation performance. For example, we could consider fisheries
managers, conservation agencies, or First Nations groups as investors.
We defined asset value as the abundance of returning salmon in each
stream and value of the portfolio as the overall metapopulation
abundance. In this scenario, the equivalent to financial rate of return
is the generation-to-generation rate of change of metapopulation
abundance. We defined the financial asset investment weights as the
capacity of the stream populations --- specifically the unfinished
equilibrium stock size --- since maintaining or restoring habitat
requires money, time, and resources.

@@

The salmon metapopulation dynamics in our simulation were governed by a
spawner-return relationship with stochastic demographic noise and by
straying between populations.

@@

\noindent where $t$ represents a generation time, $i$ represents a
population, $R$ is the number of returns, $S$ is the number of spawners,
$a$ is the productivity parameter (which can vary with the environment),
and $b$ is the density-dependent term (which is used as the asset
weights in the portfolios). The term $w_{ti}$ represents first-order
autocorrelated error. Formally, $w_{ti} = w_{ti-1} \rho_w + v_{ti}$,
where $v_{ti}$ represents independent and normally distributed error
with mean 0 and standard deviation of $\sigma_v$. The parameter $\rho_w$
represents the correlation between residuals from subsequent
generations.

@@

\begin{equation}
  e_t = \alpha_e + \beta_e t
\end{equation}

\noindent where $\alpha_e$ represents the intercept and $\beta_e$ the
slope. To maintain a balanced response, we centered the trend so that
midway through the simulation (after any burn-in period) the
environmental value was at the mean environmental tolerance.

@@

\subsubsection{Spatial conservation strategies}

@@

\subsubsection{How many populations to conserve?}

@@

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  Spatial conservation

  \begin{itemize}
  \itemsep1pt\parskip0pt\parsep0pt
  \item
    Short-term environment

    \begin{itemize}
    \itemsep1pt\parskip0pt\parsep0pt
    \item
      Response diversity conservation affects risk properties of
      portfolio
    \item
      Nearly the same portfolio whether you invest in the most stable or
      the full range of responses
    \item
      Local demographic extinction may be different
    \item
      You can see this for an example metapopulation time series: c, e:
      generation-to-generation fluctuations dampened by response
      diversity - a ying for every yang
    \end{itemize}
  \item
    Long-term environment

    \begin{itemize}
    \itemsep1pt\parskip0pt\parsep0pt
    \item
      Response diversity conservation affects return properties of
      portfolio
    \item
      Depends whether you ``get it right'' in choosing the right
      populations to conserve
    \item
      Notably, those that are most productive at first are no longer so
      later
    \item
      You can see what's happening in the example time series: d, f: You
      get it right or wrong in the long run. Response diversity means
      that when one does poorly another is doing well, longterm rate of
      return is stable
    \end{itemize}
  \end{itemize}
\item
  Number

  \begin{itemize}
  \itemsep1pt\parskip0pt\parsep0pt
  \item
    Short-term environment

    \begin{itemize}
    \itemsep1pt\parskip0pt\parsep0pt
    \item
      More is usually better for risk (give ratios)
    \item
      The unknown response diversity causes a risk spread
    \item
      Example time series: Fig 4c: Can see the stabilizing effect of
      number on the rate of change
    \item
      Example time series Fig 4d: The way we've set it up, this also
      affects the absolute abundance - a double whammy
    \end{itemize}
  \item
    Long-term environment

    \begin{itemize}
    \itemsep1pt\parskip0pt\parsep0pt
    \item
      The unknown response diversity causes a spread in rate of return -
      some cases do well, some do poorly depending on the selection of
      traits
    \item
      More is also usually better in terms of risk
    \item
      More dampens demographic stochasticity
    \item
      More likely to end up with response diversity?
    \end{itemize}
  \end{itemize}
\end{itemize}

Figure \ref{f:sp-mv}

Figure \ref{f:n-mv}

Figure \ref{f:eg-n-linear} Figure \ref{f:eg-n-arma} Figure
\ref{f:eg-sp-arma} Figure \ref{f:eg-sp-linear}

@@

\textit{Population dynamics parameters}                                                             &                       & \\
Stock-recruit residual standard deviation (on log scale)                                   & $\sigma_v$            & 0.30  \\
AR(1) serial correlation of stock-recruit residuals                                        & $\rho_w$              & 0.40  \\
Fraction of fish that stray from natal streams                                             & $f_{\mathrm{stray}}$  & 0.02  \\
Exponential rate of decay of straying with distance                                        & $m$                   & 0.3  \\

\noalign{\vskip 3mm} 
\textit{Environmental parameters}                                                                   &                       & \\
Width of environmental-tolerance curves for populations $i$ 1 to $n$                       & $W_i$                 & seq(0.05, 0.02, 0.02 0.05, length = n pop)\\
Optimum environmental value for populations $i$ 1 to $n$                                   & $e_i^{\mathrm{opt}}$  & seq(13, 19, length = n pop)\\
Area under each environmental-tolerance curve in environmental units                       & $A$                   & 30\\
AR(1) autocorrelation of the environmental signal for the short-term fluctuation scenarios & $\rho_e$              & 0.1 \\
Standard deviation of environmental noise for the short-term fluctuation scenarios         & $\sigma_d$            & 2 \\

>>>>

\begin{itemize}
\item
  Prioritizing metapopulation conservation through the lens of portfolio
  theory
\item
  Portfolio optimization reveals\ldots{}
\item
  Portfolio theory reveals conservation strategies that\ldots{}
\item
  Intelligent tinkering with ecological portfolios: patterns of
  environmental response diversity drive salmon conservation priorities
\item
  Keeping every cog and wheel\ldots{}
\end{itemize}

\textbf{Ecology Letters}

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  letters: 5000 words, 6 figures or tables
\item
  ideas and perspectives: 7500 words, 10 figures (needs 300 word
  proposal)
\end{itemize}

\begin{quote}
Ecology Letters is a forum for the very rapid publication of the most
novel research in ecology, research that is not yet in the public
domain. Manuscripts relating to the ecology of all taxa, in any biome
and geographic area will be considered, and priority will be given to
those papers exploring or testing clearly stated hypotheses. The journal
publishes concise papers that merit urgent publication by virtue of
their originality, general interest and their contribution to new
developments in ecology. We discourage purely descriptive papers and
those merely confirming or extending results of previous work.
\end{quote}

\textbf{Ecological Applications}

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  60 pages including everything
\end{itemize}

\begin{quote}
Ecological Applications is concerned broadly with the applications of
ecological science to environmental problems. It publishes papers that
develop scientific principles to support environmental decision-making,
as well as papers that discuss the application of ecological concepts to
environmental issues, policy, and management. Papers may report on
experimental tests, actual applications, scientific decision support
techniques, economic analyses, social implications of environmental
issues, or other relevant topics. Statistical or experimental methods
papers that support research and applications are welcome. Papers
submitted to Ecological Applications should be accessible to both
scholars and practitioners.
\end{quote}

@@

  \begin{itemize}
  \itemsep1pt\parskip0pt\parsep0pt
  \item
    risk is usually assessed on a population-by-population basis
  \item
    for metapopulations, we can consider risk for components or risk for
    the whole
  \end{itemize}
\item
  The management of financial portfolios provides another way of
  considering risk for metapopulations.

@@

Here, we ask how a portfolio approach to management can inform the
conservation of metapopulations in a changing world. We ask two primary
questions: (1) What does portfolio optimization tell us about different
spatial approaches to prioritizing metapopulation conservation assuming
that response diversity is spatially distributed? (2) If we don't know
how response diversity is distribute, what does portfolio optimization
tell us about how many populations we should conserve? To answer these
questions, we develop a salmon metapopulation simulation in which
population-specific productivity is driven by spatially-distributed
environmental tolerance and patterns of short- and long-term climatic
change. We then implement different conservation ``rules of thumb'' that
control the population-level carrying capacities and evaluate the salmon
portfolios along risk and return axes, as a portfolio manager would. Our
findings illustrate that: (1) conserving response diversity buffers
metapopulation risk given short-term climate forcing and rate of return
given long-term climate forcing, (2) conserving more subpopulations
buffers risk regardless of the climate trend, and (3) considering
metapopulations through the lens of portfolio theory provides a useful
additional dimension through which we can evaluate conservation
strategies.

@@

In our ecological portfolios, we defined assets as stream-level
populations and the portfolios as salmon metapopulations (Table
\ref{t:port}). We use the terms \textit{stream} and \textit{populations}
interchangeably to represent the portfolio assets. We defined the
portfolio investors as the stakeholders in the fishery and
metapopulation performance. For example, we could consider fisheries
managers, conservation agencies, or First Nations groups as investors.
We defined asset value as the abundance of returning salmon in each
stream and value of the portfolio as the overall metapopulation
abundance. In this scenario, the equivalent to financial rate of return
is the generation-to-generation rate of change of metapopulation
abundance. We defined the financial asset investment weights as the
capacity of the stream populations --- specifically the unfinished
equilibrium stock size --- since maintaining or restoring habitat
requires money, time, and resources. Investment in a population
therefore represents investing in salmon habitat conservation or
reconstruction.

@@

The salmon metapopulation dynamics in our simulation were governed by a
spawner-return relationship with demographic stochasticity and by
straying between populations.

@@

\noindent where $t$ represents a generation time, $i$ represents a
population, $R$ is the number of returns, $S$ is the number of spawners,
$a$ is the productivity parameter (which can vary with the environment),
and $b$ is the density-dependent term (which is used as the asset
weights in the portfolios). The term $w_{ti}$ represents first-order
autocorrelated error. Formally, $w_{ti} = w_{ti-1} \rho_w + r_{ti}$,
where $r_{ti}$ represents independent and normally distributed error
with mean 0 and standard deviation of $\sigma_r$. The parameter $\rho_w$
represents the correlation between residuals from subsequent
generations.

@@

\begin{equation}
  e_t = \beta_e t - \overline{\beta_e t}
\end{equation}

\noindent where $\beta_e$ represents the slope. To maintain a balanced
response, we centered the trend by subtracting the mean
$\overline{\beta_e t}$ so that midway through the simulation (after any
burn-in period) the environmental value was at the mean environmental
tolerance.

@@

@@

@@

\subsection{Which populations to conserve?}

\subsubsection{Short-term environment}

Given strong short-term environmental fluctuations, conserving response
diversity buffers the risk properties of an ecological portfolio (Fig.
4a). In our simulation, the median variance of generation-to-generation
rate of change in abundance was X times lower given balanced response
diversity (full range of responses or most stable only vs.~conserving
one half or the other). In fact, even though by conserving the full
range of responses, the portfolio was comprised of warm and
cool-thriving populations that were more variable on their own, each was
balanced by an opposing population. The portfolio risk was therefore
comparable between the full range of responses and most stable only
portfolios.

We can see the mechanism behind these portfolio properties by inspecting
example population time series (Fig. 4c, d). If only the upper or lower
half of response diversity is conserved, the portfolio tends to do well
or poorly depending on the environmental conditions (Fig. 4d). This risk
is buffered with balanced response diversity (Fig. 4c).

\subsubsection{Long-term environment}

Given long-term environmental change, the choice of which populations to
conserve affects the return properties of an ecological portfolio (Fig.
4b). By conserving balance response diversity, an ecological manager is
hedging his or her bets on what will happen with the environment and how
the populations will respond. The typical return for a balanced response
diversity strategy was zero --- the metapopulation neither increased or
decreased in abundance in the long run. By conserving only the upper or
lower half of response diversity, a conservation manager is putting all
his or her eggs in one basket --- the metapopulation might do really
well through time or it might do really poorly. The example
metapopulation abundance time series (Fig. 4d, f) illustrate this
effect. By conserving response diversity, when one population is doing
poorly, another is doing well and the metapopulation abundance remains
stationary through time.

Notably, in theses simulations, if a managers invested in the
populations that were doing well at the beginning they would have had
the lowest rate-of-return portfolio in the end (purple portfolios in
Fig. 4b).

Spatial conservation strategies in the face of longterm environmental
change hinge on whether you ``get it right'' --- whether you choose just
the right populations to conserve.

\subsection{How many populations to conserve?}

Given a scenario where we don't know the distribution of
population-level response diversity, portfolio optimization informs us
about the risk buffering from maintaining multiple populations (Fig. 5).
In short, investing in more populations buffers portfolio risk.

\subsubsection{Short-term environment}

Given short-term environmental noise, conserving more populations
buffers portfolio risk while the random conservation of response
diversity creates a spread of metapopulation risk for the same number of
populations conserved (Fig. 5a). For example, a metapopulation with
eight conserved populations is X times less risky than a metapopulation
with only four. We can see this risk-buffering effect through example
metapopulations in Fig. 5c and 5d. We note that the risk-return axes of
portfolio optimization ignore the absolute-abundance dimension (Fig.
5d). As one would expect, conserving fewer populations also results in
lower-abundance metapopulations.

\subsubsection{Long-term environment}

Given long-term environmental noise, conserving more populations also
buffers portfolio risk. However, in contrast to the short-term
environmental noise scenario, the unknown response diversity creates a
spread of possible metapopulation return for the same number of
conserved populations (Fig. 5b). Here, the number of populations
conserved buffers non-systematic (i.e.~not-environmentally-driven)
stochasticity.

@@

\textit{Population dynamics parameters}                              &                       & \\
Stock-recruit residual standard deviation (on log scale)             & $\sigma_r$            & 0.30  \\
AR(1) serial correlation of stock-recruit residuals                  & $\rho_w$              & 0.40  \\
Fraction of fish that stray from natal streams                       & $f_{\mathrm{stray}}$  & 0.02  \\
Exponential rate of decay of straying with distance                  & $m$                   & 0.3  \\

\noalign{\vskip 3mm}
\textit{Environmental parameters}                                    &                       & \\
Width of environmental-tolerance curves for populations $i$ 1 to $n$ & $W_i$                 & seq(0.05, 0.02, 0.02 0.05, length = n pop)\\
Optimum environmental value for populations $i$ 1 to $n$             & $e_i^{\mathrm{opt}}$  & seq(13, 19, length = n pop)\\
Area under each environmental-tolerance curve in environmental units & $A$                   & 30\\

Standard deviation of short-term environmental fluctuations          & $\sigma_d$            & 2 \\
AR(1) autocorrelation of short-term environmental fluctuations       & $\rho_e$              & 0.1 \\
Slope of long-term environmental signal                              & $\beta_e$             & 0.114 \\
